\Large\bf Data Intensive processing with iRODS and the middleware CiGri for the Whisper project

### 2014-06-11 articles/2014/whisper/aricleEbauche/articleIrodsUserMeeting.tex
add paragraphs until jdl for result section in subsection Whisper use case

%\documentclass[a4paper,twoside]{article}
\documentclass[a4paper, 10pt]{article}
\usepackage{graphicx}
\usepackage{rfia2000}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{multicol}
\usepackage{caption}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\setlength{\parindent}{0.3cm}

@@

\begin{multicols}{2}
%\/*
\section{Plan}
\begin{enumerate}
	\item Introduction, abstract (3\S , 1 page) 
  	\item Background (7-10\S, 1-2 pages)
    \begin{enumerate}
  	  \item The Whisper project
  	  \item Size of observational data
  	  \item Size of computational data
  	  \item General Big data/ science of  universe
  	  \item Coupling scientific objectives / IT constraints
  	  \item Data grid paradigm
  	  \item Data grid in Grenoble
  	  %\item Whisper organisation
  	  \item Collaboration Whisper/IT infrastructure
  	\end{enumerate}
  	\item Software for data-intensive processing (6-12\S , 1-2 pages)
  	\begin{enumerate}
      \item General IT codes whisper
      \item Computer language and librairies
      \item Structure of codes
      \item Package for raw data
      \item Package for correlations
      \item First optimisation of correlation
      \item Second optimisation of correlation
      \item Codes for analysis of correlations
      \item Where we can run the codes
    \end{enumerate}
  \item IT infrastrcuture for grid computing (6-12\S , 1-2 pages)
    \begin{enumerate}
  	  \item Needs of coupling storage computation
  	  \item Presentation Ciment
  	  \item Coupling Irods/Cigri
  	  \item Resulting Data-Grid
      \item Presentation Irods
	  \item General Infra Irods (diagram)
	  \item Effective Nodes Irods (diagram)
  	  \item General Cigri
  	  \item Mechanism with OAR
  	  \item Mechanism resubmission/besteffort
  	  \item Description of a campaign
  	  \item Low Interaction Cigri/Irods
  	  \item Cigri V2 and V3 functionalities
  	\end{enumerate}
  \item Results (28-31\S, 6-7 pages)
  \item Discussion (7-10\S, 1-2 pages)
\end{enumerate}
%*/

@@

>>>>

%\documentclass[a4paper,twoside]{article}
\documentclass[a4paper, 10pt]{article}
\usepackage{graphicx}
\usepackage{rfia2000}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{multicol}
\usepackage{caption}
\usepackage{listings}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\setlength{\parindent}{0.3cm}

@@

\begin{multicols}{2}
%\/*
\section{Plan}
\begin{enumerate}
	\item Introduction, abstract (3\S , 1 page) 
  	\item Background (7-10\S, 1-2 pages)
    \begin{enumerate}
  	  \item The Whisper project
  	  \item Size of observational data
  	  \item Size of computational data
  	  \item General Big data/ science of  universe
  	  \item Coupling scientific objectives / IT constraints
  	  \item Data grid paradigm
  	  \item Data grid in Grenoble
  	  %\item Whisper organisation
  	  \item Collaboration Whisper/IT infrastructure
  	\end{enumerate}
  	\item Software for data-intensive processing (6-12\S , 1-2 pages)
  	\begin{enumerate}
      \item General IT codes whisper
      \item Computer language and librairies
      \item Structure of codes
      \item Package for raw data
      \item Package for correlations
      \item First optimisation of correlation
      \item Second optimisation of correlation
      \item Codes for analysis of correlations
      \item Where we can run the codes
    \end{enumerate}
  \item IT infrastrcuture for grid computing (6-12\S , 1-2 pages)
    \begin{enumerate}
  	  \item Needs of coupling storage computation
  	  \item Presentation Ciment
  	  \item Coupling Irods/Cigri
  	  \item Resulting Data-Grid
      \item Presentation Irods
	  \item General Infra Irods (diagram)
	  \item Effective Nodes Irods (diagram)
  	  \item General Cigri
  	  \item Mechanism with OAR
  	  \item Mechanism resubmission/besteffort
  	  \item Description of a campaign
  	  \item Low Interaction Cigri/Irods
  	  \item Cigri V2 and V3 functionalities
  	\end{enumerate}
  \item Results and feedback (28-31\S, 6-7 pages))
      \begin{enumerate}
	    \item Whisper Use Case
        \item Infrastructure Ciment experiences
       \end{enumerate}
\end{enumerate}
%*/

@@

Whisper is one of the projects that have made it possible to ensure that the seismic noise brought new observable.
This permits to carry out several scientific results including imaging and monitoring. 
Concerning monitoring, further studies provide news results about sligth variations of seismics waves induce by earthquakes.
Many articles are in part due to this project \footnote{see whisper.obs.ujf-grenoble.fr, rubric publication} as well as
several post-docs and PhD.

Most of the time, on the computer part, the approach with researchers is as follows.
After retrieving data from a data center or directly between person, we have to assess how this data can be processed.
According to the computing time and storage capacity, either we perform operations on a dedicated bay (also host by Ciment), or either we use the Ciment infrastructure.
It depends also on the ease of computer users and most of the time, at least the last treatments (less computationally expensive) are made locally
(Some IT codes are also provided in order to retreive results on distributed storage).

We focus now on the part of processing that use the data grid environment of Ciment.
But note before that an other aspect, and not least, is the evolution of the specification. 
Often, students and researchers have new requirements and the IT codes evolves with these specification.
We try to be as generic as possible in order to, among other things, to achieve a sufficient level of automatization.
However, sometimes we need to develop some parts specifically because of lack of time.
These IT problems of specification and development time are among the most complex to evaluate for this type of project.

A first step, for the IT part of the scientific workflow, begins by storing data and also by checking their integrity.
We also require that data be in a seismic standard formats.
If conversion is necessary, it can be a large data intensive computing and specific codes are developed.
To minimize concurrency, we have also to ensure that the data are well distributed on iRODS. 
Indeed, it happened that the data are too centralized on a resource, so that is truly diminished processing capacity.
Some python codes are dedicated to this task and can replicate or spread randomly a Collection from a set of ressources to another set.

Let us come back to the package of signal processing of Whisper (the first part of the figure \ref{whisperGeneralWorflow}).
Because each seismogram can be treated separatly it is the simplest case for data grid process. We treated for instance one year of the
Japansese seismic Network (HiNet, Tiltmeter and FNet) around of the giant 2011 Tohoku-oki earthquake (6 months before and 6 months after).
Note that first we need to convert 9 Terabytes of Japanese Data into around 20 terabytes of a standard format (here mseed or sac).
Then we try many processing for the 20 TB (filter, whitening, clipping, ...) and store the results in seismogram with a duration of one day.

As almost all seismological data, the Japanese data are identified by the dates, names of seismic station and sub-networks. Therefore
the choice of the modelization, in order to retrieve and distribute the data, follows these seismological metadatas.
More precisely, the modelization of the distribution of the computation is made by setting a subset of dates and a subset of stations 
(This corresponds also to a normal use for researchers that wants to test some processing rapidly with a subset of seismograms).
Note also that the distribution is constant with respect to transfer.

In order to get a set of seismograms from the iRODS storage  to a computational node, we add Python modules to the Whisper package of seismogram processing.
These modules contains classes that permits to retrieve a subset of seismograms (defined by the two subsets described above). More precisely, a first step is either to test directly existence of data or either building of hash tables in order to know the available seismograms.
Then we provide an iterator (in this case, a generator for Python) on available data in order to use other methods that performs the i-command
'iget'. The same approach is made for the storage of the results (with the 'iput').

Note that these commands of transfer between iRODS and computational nodes can become very difficult to achieve because of the concurrency of the queries.
To take into account this obstacle, one develops a module that provides lot of encapsulations of the i-commands (number of try, waiting time,
resubmission, with error, 'else' command,  etc...). The IT infrastructure provide also wery usefull encapsulations.

We have also to adapt our process for the grid computation with CiGri. We first define a file of parameters 'param.txt' where
each line correponds to the parameters of one job on the grid, for instance:

\begin{lstlisting}%[language=bash]

cat param.txt

traces160_0_8_0 160 0 8 0
traces160_0_8_1 160 0 8 1
...
traces160_1_8_0 160 1 8 0
traces160_1_8_1 160 1 8 1
...
\end{lstlisting}

Here we divide the dates in 160 sublists and the stations in 8 sublists. 
The line "traces160\_0\_8\_1 160 0 8 1" correponds to a job named 'traces160\_0\_8\_1' where
we take the sublist of dates of index 0 and the sublist of stations of index 1 (each sublist have the length +/-1).
(There are also other parameters for components and networks not described here.)

For each of the 1280 jobs, a script is launched by CiGri , say 'start.bash', that take for
arguments a line of the file parameter. The script, among other things, load necessary library like appropriate python, retrieve the code on iRODS,
here 'main.py' and run the code (This is a diagrammatic view).

\begin{lstlisting}%[language=bash]

cat start.bash

#!/bin/bash
set -e
...
module load python
...
DirToCompute=$1
cd DirToCompute
...
iget -rf Codes
cd Codes
...
NumberDate=$2
IndexDate=$3
NumberStation=$4
IndexStation=$5
...
python main.py NumberDate 
     IndexDate NumberStation IndexStation
...
cd
rm -rf DirToCompute
\end{lstlisting}

Then one defines the json jdl file (job description language), say 'processingSeismogram.jdl'.

\begin{lstlisting}%[language=bash]

cat processingSeismogram.jdl

{
  "name": "test_processing",
  "resources": "core=1",
  "exec_file": "$HOME/start.bash",
  "param_file": "param.txt",
  "type": "best-effort",

  "clusters": {
    "c1": {
       	"prologue": [
       	... lines of codes for the 
       	... cluster c1 first of all ],
       	"project": "whisper",
       	"walltime": "00:20:00"
       	},
    "c2": {
    	"prologue": [
        ... lines of codes for the 
        ... cluster c2 first of all ],
        "project": "whisper",
        "max_jobs": "450",
        "walltime": "00:30:00"
    	},
    ...
    }
...
}
\end{lstlisting}

Here the campaign named 'test\_processing' take only one core. It run the file 'start.bash' with the parameter
file 'param.txt' in mode besteffort. It uses the clusters named 'c1' and 'c2' for the project 'whisper' with the duration
defined by 'walltime'. We can add other parameters for each clusters such that the maximum number of
jobs runing at the same time (the variable 'max\_jobs').
